---
title: "Analyze Data"
output: pdf_document
date: "2023-04-20"
author: "Dylan Wu"
---

## Analysis of Publications

```{r, connectToDb}
if (!hasPacman) install.packages("pacman")
pacman::p_load(RMySQL)

sqlite_conn <- dbConnect(RSQLite::SQLite(), fpath)
mysql_conn <- dbConnect(
  RMySQL::MySQL(),
  db = 'cs5200_db',
  user = 'admin',
  port = 3306,
  host = '***',
  password = '****'
)
```



### Top five journals with the most articles published in them for the time period

The most articles appeared in the journals "The Journal of pharmacy and pharmacology", "Biochimica et biophysica acta",
"The Journal of biological chemistry", "Annales de l'anesthesiologie francaise", "Biochemistry", ranging from 1036 articles
in the entire period of years 1975-1978 for "pharmacy and pharmacology" to "Biochemistry" at 375 total articles.
```{r, getTop5Journals}
res = dbGetQuery(mysql_conn, "
                    SELECT DISTINCT journal_title, articles_total 
                      FROM JournalFacts
                    ORDER BY articles_total DESC LIMIT 5
                  ")
print(res)
```

### Articles for each journal per quarter for each year

```{r getArticlesPerYearPerQtr}
res = dbGetQuery(mysql_conn, "
          SELECT
            	journal_title,
            	pub_year,
            	articles_year,
            	articles_q1,
            	articles_q2,
            	articles_q3,
            	articles_q4
          FROM
	            JournalFacts 
	        LIMIT 8
")
res

```

We can view the above data more intuitively by plotting two journals in the same plot. We differentiate between the two journals 
by styling one as dashed, and we differentiate the years by color-coding them. In this way, we can see how each journal's number
of articles change each quarter for all four years. As we can see, journal "Biochemical and biophysical" has much higher numbers 
than "Biochemical" all years except the first three quarters of 1975. Interestingly, the number of articles appearing in "Biochemical and biophysical"
increased drastically in quarter 4, which is significantly greater number of articles, at 37 articles, than all proceeding years for the journal.

```{r visualizeData}

new_df = data.frame(q1=res$articles_q1, q2=res$articles_q2, q3=res$articles_q3, q4=res$articles_q4)
years = c(1975, 1976, 1977, 1978)
colors = c("darkred", "orange", "darkgreen", "blue")

quarter_range = 1:4
plot(quarter_range, new_df[i,], ylab="Articles", xlab="Quarter", type="n", xlim=c(1,4), ylim=c(1,55), main="Articles per Quarter", xaxt = "n")
# referenced: https://stat.ethz.ch/pipermail/r-help/2004-February/046241.html; to use integer x-values (using xaxt="n" and following line)
# Note: I also used chatGPT to learn how to create plots/lines/legends in R
axis(1, at = 1:4)

# add all year-lines for both journals
abline(v = c(1,2,3,4), lty="dashed")
for (i in quarter_range) {
  lines(quarter_range, new_df[i,], lty="dashed", lwd=3, col=colors[i])
  lines(quarter_range, new_df[i+4,], lwd=3, col=colors[i])
}

# add legends
legend("topright", legend=years, col=colors, lwd=2)
legend("topleft", legend=res[c(1, 5), 'journal_title'], lty=c(2, 1), lwd=2)

```


```{r, disconnectDb}
dbDisconnect(mysql_conn)
```

